---
title: Connecting R with D3 for dynamic graphics, to explore multivariate data with tours
author:
  - name: Michael Kipp
    affiliation: Monash University
    address:
    - Department of Econometrics and Business Statistics
    email:  mkipp271@gmail.com
  - name: Dianne Cook
    affiliation: Monash University
    address:
    - Department of Econometrics and Business Statistics
    email:  dicook@monash.edu 
  - name: Ursula Laa
    affiliation: Monash University
    address:
    - Department of Physics
    email:  ursula.laa@monash.edu
bibliography: references.bib
link-citations: true
fig_caption: yes
abstract: >
  The tourr package in R (REF) has several algorithms and displays for showing multivariate data as a sequence of low-dimensional projections. It can display as a movie but has no capacity for interaction, such as stop/go, change tour type, drop/add variables. The tourrGui (REF) package provides these sorts of controls, but the interface is programmed with the dated RGtk2 (REF) package. This work explores using custom messages to pass data from R to D3 (REF) for viewing, using the shiny framework.  
preamble: >
  % Any extra latex you need in the preamble
output: rticles::rjournal_article
---

```{r cache=FALSE, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
#packages:
library(bookdown)
library(knitcitations)
library(RefManageR)
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, kfigr.link=TRUE, kfigr.prefix=TRUE, cache=TRUE, fig.env = TRUE, fig.cap=TRUE, fig.height=3)

 
options("citation_format" = "pandoc")

BibOptions(check.entries = FALSE, style = "markdown", bib.style = "alphabetic", cite.style = 'alphabetic')
```

## Introduction

The tour algorithm [@gt_pp;@gt_pp_mc] is a way of systematically generating and displaying projections of high-dimensional spaces in order for the viewer to examine the multivariate distribution of data. It can do this either randomly, or by picking projections judged interesting according to some criterion or index function. The tourr package [@tourr] provides the computing and display in R [@R,@ihaka:1996] to make several types of tours, grand, guided, little and local. The projection dimension can be chosen between one and the number of variables in the data. The display, though, has no capacity for interaction. The viewer can watch the tour like a movie, but not pause it and restart, or change tour type, or number of variables. 

These interactive controls were provided with the tourrGui package [@tourrGui], with was programmed with the RGtk2 package [@RGtk2]. This is not the toolkit of choice today, and has been superceded with primarily web-capable tools, like shiny [@shiny]. To display dynamic graphics though, still takes some work. This work explores the use of D3 [@2011-d3] as the display engine.


## Section title in sentence case

This section may contain a figure such as Figure \ref{figure:rlogo}.

\begin{figure}[htbp]
  \centering
  \caption{The logo of R.}
  \label{figure:rlogo}
\end{figure}

## Another section

There will likely be several sections, perhaps including code snippets, such as:

```{r}
x <- 1:10
x
```

## Summary

This file is only a basic article template. For full details of _The R Journal_ style and information on how to prepare your article for submission, see the [Instructions for Authors](https://journal.r-project.org/share/author-guide.pdf).

## References

<div id="refs"></div>
